Towards Spherical Robots For Mobile Mapping In Human Made Environments\tnoteref{mytitlenote}

### 2021-05-13 Latex/sections/registration.tex
small typos

Another consideration to speed up SDG is to adaptively recalculate $\boldsymbol\alpha$ for each iteration. 
We employ and modify ADADELTA as a technique to do so, which is described in detail in~\cite{zeiler2012adadelta}.
The main idea is the following:
It extends the SDG algorithm by two terms.
First, an exponentially decaying average of past gradients, which is recursively defined as
\begin{align}
    \vec{G}_{k+1} = \rho \vec{G}_{k} + (1 - \rho) {\nabla E}^2
\end{align}
and second, an exponentially decaying average of past changes, which is defined as
\begin{align}
    \vec{X}_{k+1} = \rho \vec{X}_{k} + (1 - \rho) {\boldsymbol\alpha \nabla E}^2
\end{align}
where $\rho \leq 1$ is a decay constant, typically close to $1$.
The root mean squared (RMS) of these quantities are
\begin{align}
    RMS[\vec{G}]_{k} = \sqrt{\vec{G}_{k} + \epsilon}
\end{align}
and 
\begin{align}
    RMS[\vec{X}]_{k} = \sqrt{\vec{X}_{k} + \epsilon}
\end{align}
where $\epsilon > 0$ is a very small constant, typically close to $0$.
It will prevent dividing by zero in the recalculation of $\boldsymbol\alpha$, which is as follows:
\begin{align}
    \boldsymbol\alpha_{k} = \frac{RMS[X]_{k-1}}{RMS[G]_{k}}
    \label{eq:adadeltaalphaupdate}
\end{align} 
For our particular application, ADADELTA behaves a little too aggressively.
Despite giving a good measure on how to adapt $\boldsymbol\alpha$, the algorithm sometimes overshoots, and convergence doesn't happen.
Therefore, we employ another scaling factor, typically not found in ADADELTA, extending eq.~\eqref{eq:adadeltaalphaupdate} to:
\begin{align}
    \boldsymbol\alpha_{k} = \boldsymbol\alpha_0 \cdot \frac{RMS[X]_{k-1}}{RMS[G]_{k}}
    \label{eq:adafinalized}
\end{align} 
where $\boldsymbol\alpha_0 $ holds the scaling factors for each dimension.

Finally, the SDG model is improved using eq.~\eqref{eq:adafinalized} and extends to 
\begin{align}
    \vec{\Pi}_{k+1} = \vec{\Pi}_{k} - \boldsymbol\alpha_0  \frac{RMS[X]_{k-1}}{RMS[G]_{k}} \cdot \nabla E
\end{align}
Using this algorithm once after finding correspondences from points to planes leads to convergence to a local minimum, which often isn't an optimal solution.
Even if  we increase the number of iterations dramatically, no better solution than the local minimum is found.
That is unless you consider updating the correspondence model after $i$ iterations of gradient descent.
Re-assigning point-to-plane correspondences this way $k$ times, and if $k$ is chosen large enough, it leads to an optimal solution after maximum $n = k\cdot i$ iterations of gradient descent.

@@

Secondly, we employ a continuous iteration, where the scans are processed sequentially. 
This is especially useful for mobile systems, where pose error accumulates due to increasing tracking uncertainties.
The assumption is that the error in one scan is also present in the following scan, plus some unknown new error.
We eliminate the error from scan $m$ which is also present in scan $m+1$ by applying the pose change from scan $m$, $\vec{\Pi}_{n,m} - \vec{\Pi}_{0,m}$ after $n$ gradient descent iterations, to scan $m+1$, before restarting gradient descent.

>>>>

Another consideration to speed up SDG is to adaptively recalculate $\boldsymbol\alpha$ for each iteration. 
We employ and modify ADADELTA as a technique to do so, which is described in detail in~\cite{zeiler2012adadelta}.
The main idea is the following:
It extends the SDG algorithm by two terms.
First, an exponentially decaying average of past gradients, which is recursively defined as
\begin{align}
    \vec{G}_{k+1} = \rho \vec{G}_{k} + (1 - \rho) {\nabla E}^2
\end{align}
and second, an exponentially decaying average of past changes, which is defined as
\begin{align}
    \vec{X}_{k+1} = \rho \vec{X}_{k} + (1 - \rho) {\boldsymbol\alpha \nabla E}^2
\end{align}
where $\rho \leq 1$ is a decay constant, typically close to $1$.
The root mean squared (RMS) of these quantities are
\begin{align}
    RMS[\vec{G}]_{k} = \sqrt{\vec{G}_{k} + \epsilon}
\end{align}
and 
\begin{align}
    RMS[\vec{X}]_{k} = \sqrt{\vec{X}_{k} + \epsilon}
\end{align}
where $\epsilon > 0$ is a very small constant, typically close to $0$.
It will prevent dividing by zero in the recalculation of $\boldsymbol\alpha$, which is as follows:
\begin{align}
    \boldsymbol\alpha_{k} = \frac{RMS[X]_{k-1}}{RMS[G]_{k}}
    \label{eq:adadeltaalphaupdate}
\end{align} 
For our particular application, ADADELTA behaves a little too aggressively.
Despite giving a good measure on how to adapt $\boldsymbol\alpha$, the algorithm sometimes overshoots, and does not converge.
Therefore, we employ another scaling factor, typically not found in ADADELTA, extending eq.~\eqref{eq:adadeltaalphaupdate} to:
\begin{align}
    \boldsymbol\alpha_{k} = \boldsymbol\alpha_0 \cdot \frac{RMS[X]_{k-1}}{RMS[G]_{k}}
    \label{eq:adafinalized}
\end{align} 
where $\boldsymbol\alpha_0 $ holds the scaling factors for each dimension.

Finally, the SDG model is improved using eq.~\eqref{eq:adafinalized} and extends to 
\begin{align}
    \vec{\Pi}_{k+1} = \vec{\Pi}_{k} - \boldsymbol\alpha_0  \frac{RMS[X]_{k-1}}{RMS[G]_{k}} \cdot \nabla E
\end{align}
Using this algorithm once after finding correspondences from points to planes leads to convergence to a local minimum, which is often not an optimal solution.
Even if we increase the number of iterations dramatically, no better solution than the local minimum is found.
That is unless you consider updating the correspondence model after $i$ iterations of gradient descent.
Re-assigning point-to-plane correspondences this way $k$ times, with a large enough $k$, leads to an optimal solution after maximum $n = k\cdot i$ iterations of gradient descent.

@@

Secondly, we employ a continuous iteration, where the scans are processed sequentially. 
This is especially useful for mobile systems, where pose errors accumulate due to increasing tracking uncertainties.
The assumption is that the error in one scan is also present in the following scan, plus some unknown new error.
We eliminate the error from scan $m$ which is also present in scan $m+1$ by applying the pose change from scan $m$, $\vec{\Pi}_{n,m} - \vec{\Pi}_{0,m}$ after $n$ gradient descent iterations, to scan $m+1$, before restarting gradient descent.

### 2021-05-13 Latex/sections/results.tex
small typos

Before the registration, the corridor is only represented acceptably in the front part. 
The further into the corridor, i.e., the longer the robot accumulates error, the more imprecise the date becomes. 
Finally, we see that a large portion of the data even exceeds either the \SI{1}{\meter} error, contributing to the large spike in the histogram or even the \SI{2}{\meter} error limit; thus, being cut from the representation. 
After registration, we see that, qualitatively, the ideal corridor was nearly restored from the noisy data. 
In particular a very large portion (95\%) have less than \SI{17}{\centi\meter} distances.
Further, the square and straight shape of the corridor is restored well, and especially the large spike at errors of greater than \SI{1}{m} is removed. 
Any such errors tend to occur at the back and the front of the corridor where the measured range is the largest hence has the largest contribution of the range error.

@@

Figure~\ref{fig:cylon-corrected} shows the results obtained before and after employing the plane base registration on the dataset acquired by the floating sphere.
We combine ten scans into one meta-scan, which is then globally registered.
This does not only speed up convergence due to the proportional effect on the error function but also decreases the risk of transforming the scan in a senseless way.
Transformations like these happen when incorrectly labeled points have a huge effect on the error gradient, which is especially the case when considering only a few points.

>>>>

Before the registration, the corridor is only represented acceptably in the front part. 
The further into the corridor, i.e., the longer the robot accumulates errors, the more imprecise the data becomes. 
Finally, we see that a large portion of the data even exceeds either the \SI{1}{\meter} error, contributing to the large spike in the histogram or even the \SI{2}{\meter} error limit; thus, being cut from the representation. 
After registration, we see that, qualitatively, the ideal corridor was nearly restored from the noisy data. 
In particular a very large portion (95\%) have less than \SI{17}{\centi\meter} distances.
Further, the square and straight shape of the corridor is restored well, and especially the large spike at errors of greater than \SI{1}{m} is removed. 
Any such errors tend to occur at the back and the front of the corridor where the measured range is the largest hence has the largest contribution of the range error.

@@

Figure~\ref{fig:cylon-corrected} shows the results obtained before and after employing the plane based registration on the dataset acquired by the floating sphere.
We combine ten scans into one meta-scan, which is then globally registered.
This does not only speed up convergence due to the proportional effect on the error function but also decreases the risk of transforming the scan in a senseless way.
Transformations like these happen when incorrectly labeled points have a huge effect on the error gradient, which is especially the case when considering only a few points.
